---
title: Blog Post 7
author: R package build
date: '2022-10-23'
slug: []
categories: []
tags: []
---

Welcome to my 2022 Midterm blog series. I’m Kaela Ellis, a junior at Harvard College studying Government. This semester I am taking Gov 1347: Election Analytics taught by Professor Ryan Enos, which has led to the creation of this blog. I will use this blog series to predict the 2022 Midterm election. The series will be updated weekly with new predictive measures and analyses of past elections.

This week I examined how shocks affect the polls, and if these shocks can have lasting effects on the election. In Christopher Wlezien and Robert Erikson’s “The Timeline of Presidential Election Campaigns”, they explain the difference between permanent bumps and temporary bounces. Temporary bounces are events that have an immediate effect on the polls, but their long-term effects wain over time to the extent that they do not affect the election. This compares to permanent bumps, which are events that affect the election. However, if temporary bounces occur close to election day, they affect the election. They also propose an online processing model, under which, the campaign effect cumulates, rather than decay as the election goes on. Thus, each shock has a permanent contribution to the series. However, as both parties experience shocks, the effects of a single shock typing the scale of the election are unlikely. 

There is the controversial shock described by Christopher Achen and Larry Bartels in “Democracy for Realists: Why Elections do not Produce Responsive Government”. In this, they demonstrate how the shock of the New Jersey shark attacks in 1916 led to Wilson losing three percentage points in beach counties. They ask: how can shark attacks affect elections? They conclude that voters engage in blind retrospection, meaning that they are incapable of assessing causation and mechanically translate decreased welfare toward electoral punishment. They support their blind retrospection theory with other examples, such as Al Gore losing 2.8M votes because states were too wet or dry. 

In this post, I will take the national shock of the Supreme Court granting certiorari in two affirmative action cases: Students for Fair Admissions v. President and Fellows of Harvard College and Students for Fair Admissions v. the University of North Carolina. The Supreme Court has upheld affirmative action in the past, but its decision to review these cases indicates that there may be a change. Since conservatives are against affirmative action, I anticipate that the Republicans will support the Supreme Court granting certiorari. However, there will likely not be any long-term impact witnessed in the polls, if there is any impact at all. 

To evaluate the effect of this shock, I analyzed the number of times that affirmative action supreme court cases were mentioned in the New York Times each week and compared this to the generic ballot. 


```{r, message = FALSE, echo=FALSE}
library(tidyverse)
library(janitor)
library(glmnet)
library(sf)
library(plotly)
library(usmap)
library(rmapshaper)
library(blogdown)
library(gridExtra)
library(stargazer)
library(lubridate)
library(dplyr)
#library(caret)
library(leaps)
library(ggthemes)
#library(usdata)
#library(gt)
#library(gtsummary)
library(cowplot)
```

```{r, message = FALSE, echo=FALSE}
Scrape_df <- readRDS("~/Desktop/Gov1347/Gov-1347/Week 7/dobbs_2022 (2).RDS")
polls <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 3/dist_polls_2018-2022.csv")

Scrape_df <- Scrape_df %>% 
  mutate(publ_date = substr(pub_date, 1, 10)) %>% 
  mutate(week = strftime(publ_date, format = "%V"))

# plot
#Scrape_df %>% group_by(week) %>%  dplyr::summarize(count = n()) %>% plot(Scrape_df$week, Scrape_df$count, type="l", col = "darkred", main = "Affirmitive Action Article Mentions")



#Scrape_df %>% 
#  group_by(week) %>% 
#  dplyr::summarize(count = n()) %>% 
#  ggplot(aes(week, count, group = 1, color = count)) +
#    geom_line() +
#    labs(y = "Article Count", x = "Week",
#         title = "NYT Articles mentioning Affirmitive Action Supreme Court decision in 2022",
#         color = "")
```

```{r, message = FALSE, echo=FALSE}

# how about visualization by week
# extract raw date
Scrape_df <- Scrape_df %>% 
  mutate(publ_date = substr(pub_date, 1, 10))

# mutate week variable
Scrape_df <- Scrape_df %>% 
  mutate(Scrape_df = strftime(publ_date, format = "%V"))

# plot
Scrape_df %>% 
  group_by(week) %>% 
  dplyr::summarize(count = n()) %>% 
  ggplot(aes(week, count, group = 1, color = count)) +
    geom_line() +
    labs(y = "Article Count", x = "Week",
         title = "Affirmitive Action Article Mentions",
         color = "") + # now add line for when decision was leaked
      geom_segment(x=3.5, xend=3.5,y=0,yend=37, lty=2, color="purple", alpha=0.4) +
      annotate("text", x=3.5, y=35, label="granted certiorari", size=3) 

#####
# now compare this to generic ballot
X538_generic_ballot_averages_2018_2022 <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 3/538_generic_ballot_averages_2018-2022.csv")
gb <- X538_generic_ballot_averages_2018_2022 
# convert dat
gb <-gb %>%
  mutate(date_ = mdy(date)) %>%
  mutate(year = substr(date_, 1, 4)) %>%
  filter(year == 2022) %>%
  mutate(week = strftime(date_, format = "%V")) # Jan 1 looks weird 

# get avg by party and week
dem <- gb %>%
  filter(candidate == 'Democrats')
library(plyr)
x <- ddply(dem, .(week), function(z) mean(z$pct_estimate))
x$candidate <- c('Democrats')
x$avg_dem <- x$V1
x <- x %>%
  select(-V1)
x$avg_dem <-  round(x$avg_dem , digits = 1)


rep <- gb %>%
  filter(candidate == 'Republicans')
y <- ddply(rep, .(week), function(z) mean(z$pct_estimate))
y$candidate <- c('Republicans')
y$avg_rep <- y$V1
y <- y %>%
  select(-V1) 
y$avg_rep <-  round(y$avg_rep, digits = 1)

#put all data frames into list
df_list <- list(gb, x, y)   
#merge all data frames together
polls_df <- df_list %>% reduce(full_join, by=c("candidate", "week"))
# remove NAs
polls_df[] <-  t(apply(polls_df, 1, function(x) c(x[!is.na(x)], x[is.na(x)])))
polls_df <- polls_df %>%
  select(-avg_rep) 
polls_df$avg_support <- polls_df$avg_dem
polls_df <- polls_df %>%
  select(-avg_dem) 
# keep only unique dates
polls_df <- polls_df %>%
  distinct(cycle, week, date_, avg_support, candidate) %>%
  filter(week != 52)

```
The graph above shows the number of articles that mentioned affirmative action supreme court cases each week. After the supreme court agreed to review the cases, the number of the article mentions increased to an all-time high of approximately 18 article mentions a week. This spike lasted returned to mid-levels of article mentions of around 5 articles a week for the remainder of 2022. 


```{r, message = FALSE, echo=FALSE}
polls_df %>%
  #group_by(candidate == 'Democrats') %>%
  #mutate(date_ = as.Date(date_)) %>%
  ggplot(aes(x = week, y = avg_support,
             colour = candidate)) +
  geom_line(aes(group=candidate), size = 0.3) + scale_color_manual(values = c("blue", "red")) + geom_point(size = 0.3) +
    #scale_x_date(date_labels = "%b, %Y") +
  ylab("Average Support") + xlab("Week") +
  labs(title = "2022 Generic Ballot") +
    theme_classic() + 
  # now add line for when decision was leaked and released
      geom_segment(x=3.5, xend=3.5,y=0,yend=33, lty=2, color="black", alpha=0.4) +
      annotate("text", x=3.5, y=31, label="granted certiorari", size=3)
```

The graph above shows the generic ballot in 2022. It seems that the supreme court granting certiorari had no effect on the generic ballot. The Republicans and the Democrats were both experiencing an upward trend in the polls, and after the supreme court granted certiorari, this trend continued. Therefore, I would not classify this as a shock, and will not use it in my election forecast. Thus, I return to my past prediction that Republicans will take the House majority with 221 seats. 