---
title: Post-election Campaign Narrative
author: Kaela Ellis
date: '2022-12-06'
slug: []
categories: []
tags: []
---
Welcome to my 2022 Midterm blog series. I’m Kaela Ellis, a junior at Harvard College studying Government. This semester I am taking Gov 1347: Election Analytics taught by Professor Ryan Enos, which has led to the creation of this blog. In this blog post, I will analyze a district’s election and why it deviated from the forecasts. 

Back in September, I chose to track the congressional election in Colorado’s 8th district. I thought that this would be an interesting election to follow because it is a brand-new district that was created after the 2020 census. This is the first election that this district is participating in, and therefore the district has no incumbents. This removes all of the incumbency advantages and potential incumbent pork-barrel spending plays. Additionally, there is a lack of certainty about how this electorate voted in past elections. Experts speculate that if Colorado’s 8th district existed in 2016, Trumps would’ve won the district by 2%, and in 2020 Biden would have won it by 4% points. However, experts (FiveThirtyEight) also predicted that Republican Barbara Kirkmeyer would defeat Democrat Yadira Caraveo with a certainty of victory in 91 in 100 simulations. By percent it was predicted that Republicans would take the seat by 1.3% points; but, in reality, Caraveo defeated Kirkmeyer 48.4% to 47.7%. As seen in the chart below, polls also incorrectly predicted the election. In this post, I will examine why the forecasts were wrong and how Caraveo won. 

```{r, message = FALSE, echo=FALSE}
library(tidyverse)
library(janitor)
library(glmnet)
library(sf)
library(plotly)
library(usmap)
library(rmapshaper)
library(blogdown)
library(gridExtra)
library(stargazer)
library(lubridate)
library(dplyr)
#library(caret)
library(leaps)
library(ggthemes)
#library(usdata)
#library(gt)
#library(gtsummary)
library(cowplot)
dist.polls <- read.csv("~/Desktop/Gov1347/Gov-1347/Week 3/dist_polls_2018-2022.csv")
```

```{r, message = FALSE, echo=FALSE}
dist.polls <-  subset(dist.polls, state %in% c("Colorado"))
dist.polls <- subset(dist.polls, seat_name %in% c("District 8"))
dist.polls %>% select(start_date, end_date, pollster, sample_size, partisan, candidate_name, answer, pct)
```

First, let’s examine some facts about the district and the candidates running. Colorado’s 8th is north of Denver in the working-class part of the Denver metro. It has the smallest share of active registered voters of any of Colorado’s congressional districts. It also has the state’s largest percentage of Hispanic residents at 39% of the district’s population. Democrats had an advantage when it came to fundraising with $2.7M to Republicans’ $1.1M. 

The two candidates ran very different campaigns, and before I examine their campaigns, I will first outline the arguments by Lynn Vavreck in “The Message Matters” on what is the right campaign for each candidate. In Vavreck’s novel, she explains that candidates can either be clarifying or insurgents. Clarifying candidates are candidates that the current economic conditions benefit. This is usually because either they are of the incumbent party when the economy is good or the challenging party when the economy is bad. In this case, Republicans are the clarifying candidate because they are the challenging party when the economy is bad – inflation rates are high and there is speculation that we are on the edge before a financial crisis. On the other hand, insurgent candidates are not benefited from economic conditions –  in this case, Democrats. Vavreck argues your classification as either a clarifying or insurgent candidate should determine the type of campaign that the respective candidate runs. Clarifying candidates should talk about the economy more than any other issue, and if they fail to do so they will only win ⅖ elections. Meanwhile, insurgent candidates should refocus the election away from the economy. They should discuss an issue that they’re close to most voters on and the clarifying candidate is constrained to an unpopular opinion on. Vavreck argues that if insurgent candidates fully meet this theory’s criteria, they can greatly boost their electoral outcomes. Vavreck gives many examples of successful and unsuccessful insurgents. One successful insurgent was JFK who campaigned against Nixon on the Soviet issue. Nixon was constrained on this issue, and Kennedy won the election. 

So by Vavreck’s logic, Kirkmeyer as the clarifying candidate should have run with the economy being the primary issue, while Caraveo should have run with an issue aside from the economy being the primary issue. Both candidates did precisely this. Democrat Dr. Caraveo is a current practicing doctor, and made issues relating to health her platform. Two of her major issues were abortion and medicare. Caraveo, a vocal pro-choice advocate, constrained Kirkmeyer to her pro-life position, an unpopular opinion. Caraveo also made this issue personal as she discusses how Kirkmeyer would criminalize physicians who performed abortions. In Caraveo’s ads, she says “Kirkmeyer has called for a total ban on abortion including cases of rape and incest.” Here,  Caraveo choose an issue that is highly salient to the electorate, that she holds the popular opinion on, and that Kirkmeyer is constrained on. Caraveo also takes on the issue of medicare, arguing that Kirkmeyer would end medicare for seniors. Again, Caraveo holds the popular opinion, and, as a doctor, this is a personal issue to her. Overall, Caraveo’s campaign conforms to Vavreck’s theory. Caraveo refocuses the election away from the economy and onto health issues, such as abortion and medicare. 

Republican Kirkmeyer also conforms to Vavreck’s theory. More than anything Kirkmeyer talks about the economy. In one of her ads, she says “I fought the big spenders, I balanced the county budget, and zeroed out the debt”. Kirkmeyer represents Colorado's only large, debt-free county. However, I argue that Kirkmeyer could have spent less time arguing against the legalization of fentanyl. This is not a highly salient issue, and she should have used this time to continue increasing the economy’s saliencey. 

Forecasts incorrectly predicted the election. Most forecasters predicted a Republican win with high certainty. According to Ballotpedia, forecasts predicted that Republicans would win by 1.3% points, but they actually lost by 0.7%, meaning that the forecasts were 2% points off. This is a fairly large difference, but forecasters also had limited information because Colorado’s 8th was a new district. The more significant mistake was that forecasters predicting with high certainty. FiveThirtyEight predicted that Republicans would win with a certainty of 91 out of 100 when Democrats won. 

Since both campaigns conformed to Vavreck’s theory, it was not that these candidates did not run the right kind of campaign. I believe that Kirkmeyer had a few different disadvantages. First, Caraveo’s family immigrated to Colorado from Mexico, indicating that she likely has a leg-up with the Hispanic community, which makes up 39% of the district’s population. This is an inherent disadvantage for Kirkmeyer. Secondly, Caraveo spent more than double Kirkmeyer, spending $2.7M to her $1.1M. Lastly, while Kirkmeyer ran the right type of campaign, she could have run it better. Many of her ads are about her position against fentanyl’s legalization, which is not a highly salient issue. She could have spent more time advertising her economic success and how she will help the economy. I think that these three issues contributed to Kirkmeyer’s downfall in the election. 

Sources:
Lynn Vavreck. The message matters: the economy and presidential campaigns. Princeton University Press, 2009.
https://coloradosun.com/2021/11/22/8th-congressional-district-colorado/
https://ballotpedia.org/Colorado%27s_8th_Congressional_District_election,_2022
https://www.kirkmeyerforcongress.com/about 
https://www.caraveoforcongress.com/about
https://www.politico.com/2022-election/race-forecasts-ratings-and-predictions/colorado/house/district-08/ 
