---
title: Blog_Post_6
author: Kaela Ellis
date: '2022-10-20'
slug: []
categories: []
tags: []
---

Welcome to my 2022 Midterm blog series. I’m Kaela Ellis, a junior at Harvard College studying Government. This semester I am taking Gov 1347: Election Analytics taught by Professor Ryan Enos, which has led to the creation of this blog. I will use this blog series to predict the 2022 Midterm election. The series will be updated weekly with new predictive measures and analyses of past elections.

This week I decided to incorporate turnout, incumbency, and expert predictions into my district-level two-party vote share prediction. As we saw from previous weeks campaigns do three things: persuade voters to support their candidate, turnout voters to support their vote, and convince voters to donate. Campaigns are not very effective in persuading voters to support their candidate. They have greater efficacy when trying to increase voter turnout. For example, as the Enos and Fowler reading found, the Obama campaign in 2012 increased voter turnout by approximately 8 percentage points, showing that campaigns have the potential to increase turnout by 8 points. 

However, some of the readings from this week also explored the drawbacks of ground game efforts. For example Enos and Hersh point out that canvassers are more ideologically extreme and have different policy priorities than the median voter. They point out that canvassers are likely so ideologically extreme because citizens who have ideological extremes gain more utility in pushing their extreme positions. Additionally, partaking in political activism can make one more polarized. Enos and Hersh found that 73% of the mass public thought that the economy was the most important issue, while less than 40% of canvassers thought so. This demonstrates how canvassers have different policy priorities. This leads to a principal-agent problem, as candidates have limited control over volunteer canvassers. As a result of factors such as these, Enos and Hersh found that voters contacted by in-person Obama canvassers were less likely to support Obama, demonstrating a backfire effect. 

This week I used voter turnout as a proxy for turnout. I worked on some of the code for this blog post with Lucy Ding and Jude Park. We created a district-by-district forecast, including the variables: average support, turnout, GDP, and incumbency. I used these variable in the model as follows: 

I calculated district level turnout by adding the Republican vote share and adding it to the Democratic vote share, and then dividing it by the current voting age population (CVAP). While this is not necessarily the best way to predict turnout, we have the most readily available data on this. The main drawback to using this I found was that I could only find data on turnout tracking back to 2012. This is clearly very limited, as it only gives me 5 points of data to predict my forecast off of. With less points of data, the accuracy of my forecast is limited. Other methods for calculating turnout that would be more reflective of ground game efforts would be a measure of some of the factors that Gerber and Green discussed in their 2015 paper. They discussed how get out the vote efforts that emphasized one's civic duty, polling place location, and reminder of an early pledge to vote are statistically proven to increase voter turnout. Meanwhile other GOTV efforts, such as leaflets, signage, direct mail reminders, and emails, have no apparent effect. A measure of the effective GOTV methods may inform how campaign efforts can effect election turnout more accurately. However, this data is not accessible, and therefore I defaulted to using the CVAP.  

To calculate average support, I averaged the generic ballot for 52 days prior to the election. I have discussed this decision in earlier blog iterations in greater detail. I may pivot to using a different metric in my final calculation, but the general thought here was that generic ballots closer to the election are more accurate and, therefore, better predictors of the election results. 

To calculate the economic factor, I used quarter 6 to 7 difference for GDP. Again, I have discussed different metrics for determining the economic effect. This is not the best economic factor, and I may pivot in my final calculation, but my thought here was that the change in GDP from Q6 to Q7 is the most recent, and, therefore, the most salient. 

For incumbency, I calculated whether or not the Democratic party was the incumbent. I have discussed this decision in past blog iterations, but the general thought here was that there is an incumbency advantage, and to calculate its affect on the Democratic vote share I code for incumbency in a 0:1 form. 


```{r, message = FALSE, echo=FALSE, warning = FALSE}
library(tidyverse)
library(janitor)
library(glmnet)
library(sf)
library(plotly)
library(usmap)
library(rmapshaper)
library(blogdown)
library(gridExtra)
library(stargazer)
library(lubridate)
#library(caret)
library(leaps)
library(ggthemes)
#library(usdata)
#library(gt)
#library(gtsummary)
library(cowplot)

```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
incumb <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 4/incumb_dist_1948-2022 (2).csv")

historical_results <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 4/house party vote share by district 1948-2020.csv") %>% 
  clean_names()
dem_results <- historical_results %>% 
  select(race_year, state, area, dem_votes_major_percent, district_num) %>% 
  rename("year" = "race_year") %>%  rename("DISTRICT" = "district_num") %>% rename("STATENAME" = "state") %>% filter(year ==2018)
incumb <- incumb %>%  mutate(incumb = ifelse(winner_candidate == "DemCandidate",1,0))%>% filter(year > 2008) %>% select(year, st_cd_fips, incumb, winner_candidate, state, district_num) %>% mutate(year = year +2) %>% select (year, st_cd_fips, incumb, state, district_num)

```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
turnout <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 6/turnout.csv")
output <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 6/preds_df.csv")

turnout2 <- turnout %>% rename("DISTRICT" = "district_num") %>% rename("STATENAME" = "state") %>% rename("state_dists" = "st_cd_fips") %>% filter(year == 2020) %>% select(state_dists, DISTRICT, STATENAME)

output <- output %>% left_join(turnout2, by=c("state_dists"))

get_congress_map <- function(cong=114) {
  tmp_file <- tempfile()
  tmp_dir  <- tempdir()
  zp <- sprintf("https://cdmaps.polisci.ucla.edu/shp/districts114.zip",cong)
  download.file(zp, tmp_file)
  unzip(zipfile = tmp_file, exdir = tmp_dir)
  fpath <- paste(tmp_dir, sprintf("districtShapes/districts114.shp",cong), sep = "/")
  st_read(fpath)
}

cd114 <- get_congress_map(114)

cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
dem_results$DISTRICT <- as.numeric(dem_results$DISTRICT)
cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
dem_results$STATENAME <- as.character(dem_results$STATENAME)
cd114$STATENAME <- as.character(cd114$STATENAME)

cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
cd114 <- cd114 %>% left_join(dem_results, by=c("DISTRICT", "STATENAME"))

output <- output %>% 
  mutate(preds = ifelse(preds > 100,100, preds),
         preds = ifelse(preds < 0, 0, preds))

```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
output$DISTRICT <- as.numeric(output$DISTRICT)
cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
output$STATENAME <- as.character(output$STATENAME)
cd114$STATENAME <- as.character(cd114$STATENAME)

cd2022 <- cd114 %>% left_join(output, by=c("STATENAME", "DISTRICT")) %>% filter(!STATENAME %in% c("Alaska","Hawaii"))
cd2022 <- cd2022%>% st_as_sf() %>%
  st_transform("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +ellps=sphere +units=m +no_defs")
districts_simp_2022 <- rmapshaper::ms_simplify(cd2022, keep = 0.01)
fig1 <- ggplot() + 
  geom_sf(data=districts_simp_2022,aes(fill=preds),
          inherit.aes=FALSE,alpha=0.9) + 
  scale_fill_gradient2("Democratic \nVoteshare", midpoint = 50, high = "#30a2da", mid = "white", low = "#fc4f30") +
  #coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 49), expand = FALSE) +  
  theme_void() + labs(title = "2022 Prediction with Turnout") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

#Insert no turnout dataset here
no_TO <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 6/better_preds.csv")
no_TO <- no_TO %>% left_join(turnout2, by=c("state_dists"))

no_TO <- no_TO %>%  mutate(preds = ifelse(preds > 100,100, preds),
         preds = ifelse(preds < 0, 0, preds))

cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
no_TO$DISTRICT <- as.numeric(no_TO$DISTRICT)
cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
no_TO$STATENAME <- as.character(no_TO$STATENAME)
cd114$STATENAME <- as.character(cd114$STATENAME)

cd2022_no_TO <- cd114 %>% left_join(no_TO, by=c("STATENAME", "DISTRICT")) %>% filter(!STATENAME %in% c("Alaska","Hawaii"))
cd2022_no_TO <- cd2022_no_TO%>% st_as_sf() %>%
  st_transform("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +ellps=sphere +units=m +no_defs")
districts_simp_2022_no_TO <- rmapshaper::ms_simplify(cd2022_no_TO, keep = 0.01)
fig2 <- ggplot() + 
  geom_sf(data=districts_simp_2022_no_TO,aes(fill=preds),
          inherit.aes=FALSE,alpha=0.9) + 
  scale_fill_gradient2("Democratic \nVoteshare", midpoint = 50, high = "#30a2da", mid = "white", low = "#fc4f30") +
  #coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 49), expand = FALSE) +  
  theme_void() + labs(title = "2022 Prediction without Turnout") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

plot_grid(fig1, fig2, ncol=2)


comparison <- no_TO %>%  left_join(output, by=c("STATENAME", "DISTRICT")) %>% filter(!STATENAME %in% c("Alaska","Hawaii"))

comparison$preds.x <- as.numeric(comparison$preds.x)

comparison1 <- comparison %>% mutate(preds.x = ifelse(preds.x > 50,1,0)) %>% mutate(preds.y = ifelse(preds.y > 50,1,0))

comparison1$pred.diff <- comparison1$preds.x - comparison1$preds.y

cdcomp <- cd114 %>% left_join(comparison1, by=c("STATENAME", "DISTRICT")) %>% filter(!STATENAME %in% c("Alaska","Hawaii"))
cdcomp <- cdcomp%>% st_as_sf() %>%
  st_transform("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +ellps=sphere +units=m +no_defs")
districts_simp_comp <- rmapshaper::ms_simplify(cdcomp, keep = 0.01)
ggplot() + 
  geom_sf(data=districts_simp_comp,aes(fill=pred.diff),
          inherit.aes=FALSE,alpha=0.9) + 
  scale_fill_gradient2("Flipped \n States", midpoint = 0, high = "#30a2da", mid = "white", low = "#fc4f30") +
  #coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 49), expand = FALSE) +  
  theme_void() + labs(title = "States that Flipped turnout vs without turnout") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```
With the incorporation of incumbency into my model, I saw 36 states flip. So turnout had an affect on my predictions. With the turnout variable, I predict that Democrats will receive 220 seats. 

```{r, message = FALSE, echo=FALSE, warning = FALSE}
#comparison1 %>%
 # filter(preds.x == 0)
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
#Seats that flipped 2020 to 2022
#results from 2020 compared to results from 2022
no_TO <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 6/better_preds.csv")
turnout2 <- turnout %>% rename("DISTRICT" = "district_num") %>% rename("STATENAME" = "state") %>% rename("state_dists" = "st_cd_fips") %>% filter(year == 2020) %>% select(state_dists, DISTRICT, STATENAME)

no_TO <- no_TO %>% left_join(turnout2, by=c("state_dists"))
historical_results <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 4/house party vote share by district 1948-2020.csv") %>% 
  clean_names()
dem_results_2020 <- historical_results %>% 
  select(race_year, state, area, dem_votes_major_percent, district_num) %>% 
  rename("year" = "race_year") %>%  rename("DISTRICT" = "district_num") %>% rename("STATENAME" = "state") %>% filter(year == 2020) %>% distinct() %>% distinct(STATENAME,DISTRICT, .keep_all = TRUE)

dem_results_2020$DISTRICT <- as.numeric(dem_results_2020$DISTRICT)
no_TO$DISTRICT <- as.numeric(no_TO$DISTRICT)
dem_results_2020$DISTRICT <- as.numeric(dem_results_2020$DISTRICT)
no_TO$STATENAME <- as.character(no_TO$STATENAME)
dem_results_2020$STATENAME <- as.character(dem_results_2020$STATENAME)
diff.2020.2022 <- no_TO %>% left_join(dem_results_2020, by=c("STATENAME", "DISTRICT"))

diff.2020.2022 <- diff.2020.2022 %>%  mutate(preds = ifelse(preds > 50,1,0)) %>% mutate(dem_votes_major_percent = ifelse(dem_votes_major_percent > 50,1,0))
diff.2020.2022 <- as.data.frame(diff.2020.2022)
diff.2020.2022$pred.diff <- diff.2020.2022$preds - diff.2020.2022$dem_votes_major_percent

cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
diff.2020.2022$DISTRICT <- as.numeric(diff.2020.2022$DISTRICT)
cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
diff.2020.2022$STATENAME <- as.character(diff.2020.2022$STATENAME)
cd114$STATENAME <- as.character(cd114$STATENAME)

diff.2020.2022 <- cd114 %>% left_join(diff.2020.2022, by=c("STATENAME", "DISTRICT")) %>% filter(!STATENAME %in% c("Alaska","Hawaii"))
diff.2020.2022 <- diff.2020.2022%>% st_as_sf() %>%
  st_transform("+proj=laea +lat_0=45 +lon_0=-100 +x_0=0 +y_0=0 +ellps=sphere +units=m +no_defs")
districts_simp_2022_comb <- rmapshaper::ms_simplify(diff.2020.2022, keep = 0.01)
fig5 <- ggplot() + 
  geom_sf(data=districts_simp_2022_comb,aes(fill=pred.diff),
          inherit.aes=FALSE,alpha=0.9) + 
  scale_fill_gradient2("Democratic \nVoteshare", midpoint = 0, high = "#30a2da", mid = "white", low = "#fc4f30") +
  #coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 49), expand = FALSE) +  
  theme_void() + labs(title = "Seats flipped 2020 to 2022: Without Turnout") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
#as.data.frame(diff.2020.2022 %>%
 # filter(pred.diff != 0)) %>%  mutate(diff.2020.2022 = ifelse(pred.diff == -1,"R",pred.diff), diff.2020.2022 = ifelse(pred.diff == 1,"D",diff.2020.2022)) %>% select(STATENAME, DISTRICT,  diff.2020.2022)
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
#average <- c(221, 214)
#high <- c(224, 211)
#low<- c(215, 220)

#df <- as.data.frame(rbind(average, high, low))
#df <- df %>% rename("Republican" = "V1") %>%  rename("Democrat" = "V2")


#ggplot(df, aes(x = Democrat, y = c(average, high, low), fill = Democrat)) + 
 # geom_bar(stat = "identity") +
  #scale_fill_manual(values = c("#30a2da", "#fc4f30")) +
  #ylab("Seats") + xlab("Voter Turnout") +
  # ggtitle("Incumbency win rate by year") +
  #coord_flip() + 
  #theme_bw() + 
  #theme(axis.title = element_text(size=10),
    #    axis.text = element_text(size=5),
   #     strip.text.y = element_text(size = 5))

#df <- df %>%
  #group_by(year) %>% mutate(tot_n=n()) %>% ungroup() %>%
  ## this one is get numerator and calculate % by party
 # group_by(year, winner_candidate_inc) %>% summarise(p_n=n()*100/first(tot_n)) %>% ungroup() %>%
   ## filter idiosyncratic issues
#  filter(p_n != 100.000000) %>%
  ## finally, this one so we can sort the issue names
  ## by D% of issue ad-share instead of alphabetically
#  group_by(year) %>% mutate(Dp_n = ifelse(first(winner_candidate_inc) == "democrat", first(p_n), 0))

#ggplot(incumb_ed, aes(x = reorder(year, Dp_n), y = p_n, fill = winner_candidate_inc)) + 
  #geom_bar(stat = "identity") +
 # scale_fill_manual(values = c("darkgreen", "purple1")) +
 # ylab("Incumbency Win Rate") + xlab("Year") +
  # ggtitle("Incumbency win rate by year") +
#  coord_flip() + 
#  theme_bw() + 
 # theme(axis.title = element_text(size=10),
    #    axis.text = element_text(size=5),
      #  strip.text.y = element_text(size = 5)) 
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
cvap <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 6/cvap_district_2012-2020_clean.csv")
electionresults <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 6/incumb_dist_1948-2020 (3).csv")
genericballot <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 3/polls_df.csv")
gdp_q <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 2/GDP_quarterly.csv")
incumbency <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 6/incumb.csv")
```
```{r, message = FALSE, echo=FALSE, warning = FALSE}
# renaming geoid to st_cd_fips for left join
cvap <- cvap %>% 
  rename(st_cd_fips = geoid) 

# create turnout dataframe by left joining election results and cvap
turnout <-left_join(electionresults, cvap, by = c("year", "st_cd_fips")) %>% 
  select(year, state.x, district_num, winner_party, 
         RepVotes, DemVotes, DemVotesMajorPercent, st_cd_fips, cvap) %>% 
  filter(year >= 2012) %>%  # filter for 2012-2020
  rename(state = state.x) %>% 
  mutate(turnout = (RepVotes + DemVotes) / cvap * 100) # turnout column

# creating polls_d dataframe by cleaning genericballots
genericballot$poll_date <- as.Date(genericballot$poll_date, format = "%m/%d/%Y")

cleaned_polls <- genericballot %>% 
  filter(year >= 2012,
         days_until_election <= 52) %>% 
         # the closest poll to election day for 1952 was 51.5
  group_by(year, party) %>% 
  summarise(average_support = mean(support))

polls_d <- cleaned_polls %>% 
  filter(party =="D")

# adding generic ballot average to turnout
turnout <- left_join(turnout, polls_d, by = "year") %>% 
  select(year, state, district_num, winner_party, RepVotes, DemVotes,
         DemVotesMajorPercent, st_cd_fips, cvap, turnout, average_support)
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
# Create quarter change data and join all together with popular vote

# Filter out data for Q6
gdp_q6 <- gdp_q %>%
  filter(year >= 2012, 
         quarter_cycle == 6) %>% 
  rename(gdp_q6 = GDPC1)

# Filter out data for Q7
gdp_q7 <- gdp_q %>%
  filter(year >= 2012, 
         quarter_cycle == 7) %>% 
  rename(gdp_q7 = GDPC1)

# creating gdp_q_final
gdp_q_final <- left_join(gdp_q7, gdp_q6, by = "year") %>% 
  mutate(gdp_q7_q6 = (gdp_q7 - gdp_q6) / gdp_q6 * 100) %>% # percent difference
  rename(gdp_percent_difference = gdp_q7_q6) %>% 
  select(year, gdp_percent_difference)

# merging gdp_q_final and turnout
turnout <- left_join(turnout, gdp_q_final, by = "year")

# merging incumbency and turnout
turnout <- left_join(turnout, incumbency, by = c("year", "st_cd_fips")) %>% 
  select(year, state, district_num, winner_party, RepVotes, DemVotes,
         DemVotesMajorPercent, st_cd_fips, cvap, turnout, 
         average_support, gdp_percent_difference, incumb)
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
# creating 2022 dataframe
# first, add the average voter turnout in district from 2012-2020
turnout_2022 <- turnout %>% 
  group_by(st_cd_fips) %>% 
  mutate(historical_turnout_by_district = mean(turnout)) %>% 
  ungroup() %>% 
  select(year, state, district_num, st_cd_fips, historical_turnout_by_district) %>% 
  filter(year == 2012) %>% 
  select(-year)

# second, add nationwide variables
# Generic ballot: 45.6% for Democrats from FiveThirtyEight
# GDP: -0.6 (taken from Q2 change)
turnout_2022 <- turnout_2022 %>% 
  mutate(average_support = 45.6,
         gdp_percent_difference = -0.6)

# Incumbency
incumbency_2022 <- incumbency %>% 
  filter(year == 2022)

# Final 2022 data set to predict on
turnout_2022 <- left_join(turnout_2022, incumbency_2022, by = "st_cd_fips") %>% 
  select(state, district_num, st_cd_fips, historical_turnout_by_district,
         average_support, gdp_percent_difference, incumb) %>% 
  unique()

write.csv(turnout_2022, "turnout_2022.csv")
``` 

```{r, message = FALSE, echo=FALSE, warning = FALSE}
models <- list()
r2 <- c()
preds <- c()

state_dists <- unique(turnout$st_cd_fips)

for (sd in state_dists){
  # Filter for dataframe with just district sd
  temp <- turnout %>%
    filter(st_cd_fips == sd)
  
  # Fit linear model for dem two-party vote-share
  m <- lm(turnout ~ year, data = temp)
  
  # Add model to list of models
  models[[length(models)+1]] <- m
  
  # Find r^2 and add to list of r^squareds
  r2 <- c(r2, summary(m)$adj.r.squared)
  
  new_data <- data.frame(year = 2022)
  
  preds <- c(preds, predict(m, new_data))
}

# create a column with the new turnout
turnout_2022$new_turnout=preds

```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
# Load libraries
library(tidyverse)
library(janitor)
library(glmnet)
library(sf)
library(plotly)
library(usmap)
library(rmapshaper)
library(blogdown)
library(gridExtra)
library(stargazer)
library(lubridate)
#library(caret)
library(leaps)
library(ggthemes)
#library(usdata)
#library(gt)
#library(gtsummary)
library(cowplot)
```

# Data

```{r, message = FALSE, echo=FALSE, warning = FALSE}
turnout <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 6/turnout_2012_2020.csv")
turnout22 <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 6/turnout_2022.csv")
electionresults <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 6/incumb_dist_1948-2020 (3).csv")
genericballot <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 3/polls_df.csv")
gdp_q <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 2/GDP_quarterly.csv")
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
genericballot <- genericballot %>% 
  mutate(poll_data = as.Date(genericballot$poll_date, format = "%m/%d/%Y")) %>%
  filter(days_until_election <= 52) %>% 
         # the closest poll to election day for 1952 was 51.5
  group_by(year, party) %>% 
  summarise(average_support = mean(support)) %>%
  filter(party =="D")

# Filter out data for Q6
gdp_q6 <- gdp_q %>%
  filter(quarter_cycle == 6) %>% 
  rename(gdp_q6 = GDPC1)

# Filter out data for Q7
gdp_q7 <- gdp_q %>%
  filter(quarter_cycle == 7) %>% 
  rename(gdp_q7 = GDPC1)

# creating gdp_q_final
gdp_q_final <- left_join(gdp_q7, gdp_q6, by = "year") %>% 
  mutate(gdp_q7_q6 = (gdp_q7 - gdp_q6) / gdp_q6 * 100) %>% # percent difference
  rename(gdp_percent_difference = gdp_q7_q6) %>% 
  select(year, gdp_percent_difference)

incumbency <- electionresults %>%  
  mutate(incumb = ifelse(winner_candidate == "DemCandidate",1,0)) %>% 
  mutate(year = year + 2) %>% 
  select (year, st_cd_fips, incumb)


# adding generic ballot average to turnout
df <- left_join(electionresults, genericballot, by = "year") %>%
  left_join(gdp_q_final, by = "year") %>%
  left_join(incumbency, by = c("year", "st_cd_fips")) %>% 
  select(year, state, district_num, winner_party, RepVotes, DemVotes,
         DemVotesMajorPercent, st_cd_fips,
         average_support, gdp_percent_difference, incumb) %>%
  drop_na(incumb)

state_dists <- unique(turnout$st_cd_fips)
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
turnout14 <- turnout %>%
  filter(year == 2014) %>%
  select(year, turnout, st_cd_fips) %>%
  distinct() %>%
  distinct(st_cd_fips, .keep_all = TRUE)
turnout18 <- turnout %>%
  filter(year == 2018) %>%
  select(year, turnout, st_cd_fips) %>%
  distinct() %>%
  distinct(st_cd_fips, .keep_all = TRUE)
turnout22_fixed <- left_join(turnout14, turnout18, by = "st_cd_fips") %>%
  mutate(hist_avg = (turnout.x + turnout.y)/2) %>%
  select(st_cd_fips, hist_avg)
turnout22 <- left_join(turnout22, turnout22_fixed, by = "st_cd_fips") %>%
  select(-historical_turnout_by_district)
```


# Model with Turnout

```{r, message = FALSE, echo=FALSE, warning = FALSE}
# Instantiate lists
models <- list()
r2 <- c()
preds <- c()
preds_high <- c()
preds_low <- c()
coeff_support <- c()
coeff_turnout <- c()
coeff_gdp <- c()
coeff_incumb <- c()

# Loop through each district, model and predict
for (sd in state_dists){
  # Filter for dataframe with just district sd
  temp <- turnout %>%
    filter(st_cd_fips == sd)
  
  # Fit linear model for dem two-party vote-share
  m <- lm(DemVotesMajorPercent ~ average_support + turnout + gdp_percent_difference + incumb, data = temp)
  
  # Add model to list of models
  models[[length(models)+1]] <- m
  
  # Find r^2, coeffs, and add to list
  r2 <- c(r2, summary(m)$adj.r.squared)
  coeff_support <- c(coeff_support, summary(m)$coefficients[2])
  coeff_turnout <- c(coeff_turnout, summary(m)$coefficients[3])
  coeff_gdp <- c(coeff_gdp, summary(m)$coefficients[4])
  coeff_incumb <- c(coeff_incumb, summary(m)$coefficients[5])
  
  # Make a 2022 prediction and append to list
  new_data <- turnout22 %>%
    filter(st_cd_fips == sd) %>%
    rename(turnout = hist_avg)
  preds <- c(preds, predict(m, new_data))
  preds_high <- c(preds_high, predict(m, new_data %>% mutate(turnout = turnout + 8)))
  preds_low <- c(preds_low, predict(m, new_data %>% mutate(turnout = turnout - 8)))
}

output <- tibble(state_dists, preds, preds_high, preds_low, r2, coeff_gdp, coeff_incumb, coeff_support, coeff_turnout)
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
state_dists[435]
```


```{r, message = FALSE, echo=FALSE, warning = FALSE}
stargazer(m, type = "text")
```


```{r, message = FALSE, echo=FALSE, warning = FALSE}
#tbl_regression(m)
```



```{r, message = FALSE, echo=FALSE, warning = FALSE}
preds_df <- tibble(state_dists, preds, preds_high, preds_low)
#write.csv(preds_df, "preds_df.csv")
```


```{r, message = FALSE, echo=FALSE, warning = FALSE}
length(preds[preds <= 50])
length(preds[preds > 50])
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
length(preds[preds_high <= 50])
length(preds[preds_high > 50])
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
length(preds[preds_low <= 50])
length(preds[preds_low > 50])
```

```{r, message = FALSE, echo=FALSE}
#ggplot(turnout %>% filter(year == 2018), aes(x = DemVotesMajorPercent)) +
 # geom_histogram(bins = 30, color = "white") +
  #theme_few() +
  #labs(title = "Actual 2018 House Election Results",
   #    subtitle = "House Districts",
    #   x = "Democratic Two-Party Vote-Share (%)",
     #  y = "Count") +
#  ylim(0,50)
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
figmid <- ggplot(output %>% mutate(preds = ifelse(preds > 100, 100, preds),
                                        preds = ifelse(preds < 0, 0, preds)), 
       aes(x = preds)) +
  geom_histogram(bins = 30, color = "white") +
  theme_few() +
  labs(title = "2022 Predictions",
       subtitle = "Average Turnout",
       x = "Democratic Vote-Share (%)",
       y = "Count") +
  xlim(-5, 105) +
  ylim(0,50)

fighigh <- ggplot(output %>% mutate(preds_high = ifelse(preds_high > 100, 100, preds_high),
                                        preds_high = ifelse(preds_high < 0, 0, preds_high)), 
       aes(x = preds_high)) +
  geom_histogram(bins = 30, color = "white") +
  theme_few() +
  labs(title = "2022 Predictions",
       subtitle = "High Turnout",
       x = "Democratic Vote-Share (%)",
       y = "Count") +
  xlim(-5, 105) +
  ylim(0,50)

figlow <- ggplot(output %>% mutate(preds_low = ifelse(preds_low > 100, 100, preds_low),
                                        preds_low = ifelse(preds_low < 0, 0, preds_low)), 
       aes(x = preds_low)) +
  geom_histogram(bins = 30, color = "white") +
  theme_few() +
  labs(title = "2022 Prediction",
       subtitle = "Low Turnout",
       x = "Democratic Vote-Share (%)",
       y = "Count") +
  xlim(-5, 105) +
  ylim(0,50)

plot_grid(fighigh, figmid, figlow, ncol=3)

```
I then calculated how a change in the level of turnout would change the election results. I did the most extreme turnout change of 8%, since Enos and Fowler found that campaigns have the potential to effect turnout by 8 percentage points. I found that low turnout benefits Democrats, as they win the house majority with 220 seats. Then, with average turnout, Republicans take the house majority with 221 seats. Finally, with high turnout, Republicans have house majority with 224 seats. 

```{r, message = FALSE, echo=FALSE, warning = FALSE}
#output_stacked <- output %>% 
 # pivot_longer(cols = c(preds, preds_high, preds_low), names_to = "preds") %>%
#  mutate(value = ifelse(value > 100, 100, value), 
 #        value = ifelse(value < 0, 0, value))

#ggplot(output_stacked, aes(x=value, fill=preds)) +
 # geom_histogram(alpha=0.2, position='identity')
```



```{r, message = FALSE, echo=FALSE, warning = FALSE}
ggplot(output, aes(x = r2)) +
  geom_histogram(bins = 30, color = "white") +
  theme_few() +
  labs(title = "R-Squared Values for Turnout Models in Each District",
       subtitle = "House Districts",
       x = "R-Squared",
       y = "Count")
```
I then graphed the R2 values of the model for each state. Most of the values were close to 1, indicating that this is a decent model. This model shows the R2 for the forecast which includes all the variables. However, when I isolate R2 of turnout it is 0.089, which is not a statistically significant R2 value. Therefore, in later iterations of my model I likely will remove turnout from the model. 


```{r, message = FALSE, echo=FALSE, warning = FALSE}
p1 <- ggplot(output, aes(x = coeff_gdp)) +
  geom_histogram(bins = 30) + 
  theme_few() +
  labs(x = "GDP Coefficient",
       y = "Count")
p2 <- ggplot(output, aes(x = coeff_support)) +
  geom_histogram(bins = 30) +
  theme_few() +
  labs(x = "Generic Ballot Coefficient",
       y = "Count")
p3 <- ggplot(output, aes(x = coeff_turnout)) +
  geom_histogram(bins = 30) + 
  theme_few() +
  labs(x = "Turnout Coefficient",
       y = "Count")
#grid.arrange(p1, p2, p3)
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
p1 <- ggplot(output, aes(x = coeff_gdp)) +
  geom_histogram(bins = 30, color = "white") + xlim(-2.5, 2.5) + 
  geom_vline(xintercept = median(output$coeff_gdp, na.rm = TRUE),
             color = "red", size = 1) + 
  theme_few() +
  labs(x = "GDP Coefficient",
       y = "Count") +
  ylim(0, 130)

p2 <- ggplot(output, aes(x = coeff_support)) +
  geom_histogram(bins = 30, color = "white") + xlim(-10, 10) +  
  geom_vline(xintercept = median(output$coeff_support, na.rm = TRUE),
             color = "red", size = 1) + 
  theme_few() +
  labs(x = "Generic Ballot Coefficient",
       y = "Count") +
  ylim(0, 130)

p3 <- ggplot(output, aes(x = coeff_turnout)) +
  geom_histogram(bins = 30, color = "white") + xlim(-2.5, 2.5) + 
  geom_vline(xintercept = median(output$coeff_turnout, na.rm = TRUE),
             color = "red", size = 1) + 
  theme_few() +
  labs(x = "Turnout Coefficient",
       y = "Count") +
  ylim(0, 130)

p4 <- ggplot(output, aes(x = coeff_incumb)) +
  geom_histogram(bins = 30, color = "white") + xlim(-20, 20) + 
  geom_vline(xintercept = median(output$coeff_incumb, na.rm = TRUE),
             color = "red", size = 1) + 
  theme_few() +
  labs(x = "Incumbency Coefficient",
       y = "Count") +
  ylim(0, 130)

#grid.arrange(p1, p2)
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
#p3
```


```{r, message = FALSE, echo=FALSE, warning = FALSE}
models <- list()
r2 <- c()
preds <- c()
coeff_support <- c()
coeff_gdp <- c()
coeff_incumb <- c()
for (sd in state_dists){
  # Filter for dataframe with just district sd
  temp <- df %>%
    filter(year != 2022) %>%
    filter(st_cd_fips == sd)
  
  # Fit linear model for dem two-party vote-share
  m <- lm(DemVotesMajorPercent ~ average_support + gdp_percent_difference + incumb, data = temp)
  
  # Add model to list of models
  models[[length(models)+1]] <- m
  
  # Find r^2, coeffs, and add to list
  r2 <- c(r2, summary(m)$adj.r.squared)
  coeff_support <- c(coeff_support, summary(m)$coefficients[2])
  coeff_gdp <- c(coeff_gdp, summary(m)$coefficients[3])
  coeff_incumb <- c(coeff_incumb, summary(m)$coefficients[4])
  
  # Make a 2022 prediction and append to list
  new_data <- turnout22 %>%
    filter(st_cd_fips == sd) %>%
    rename(turnout = hist_avg)
  preds <- c(preds, predict(m, new_data))
}
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
#stargazer(m, type = "text")
```


```{r, message = FALSE, echo=FALSE, warning = FALSE}
better_preds <- tibble(state_dists, preds)
output_v2 <- tibble(r2, preds, coeff_support, coeff_gdp, coeff_incumb)
#write.csv(better_preds, "better_preds.csv")
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
fig10 <- ggplot(output_v2, aes(x = r2)) +
  geom_histogram(bins = 30, color = "white") +
  theme_few() +
  labs(title = "R-Squared Values for Models w/o Turnout in Each District",
       subtitle = "House Districts",
       x = "R-Squared",
       y = "Count")
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
fig20<- ggplot(output_v2 %>% mutate(preds = ifelse(preds > 100, 100, preds),
                                        preds = ifelse(preds < 0, 0, preds)), 
       aes(x = preds)) +
  geom_histogram(bins = 30, color = "white") +
  theme_few() +
  labs(title = "2022 Election Predictions - Model w/o Turnout",
       subtitle = "House Districts",
       x = "Democratic Two-Party Vote-Share (%)",
       y = "Count") +
  xlim(-5, 105) +
  ylim(0,50)
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
#p1 <- ggplot(output_v2, aes(x = coeff_gdp)) +
 # geom_histogram(bins = 30, color = "white")
  #geom_vline(xintercept = median(output_v2$coeff_gdp, na.rm = TRUE),
   #          color = "red", size = 1) + 
  #theme_few() +
#  labs(x = "GDP Coefficient",
 #      y = "Count") +
  #ylim(0, 150)

#p2 <- ggplot(output_v2, aes(x = coeff_support)) +
 # geom_histogram(bins = 30, color = "white")
  #geom_vline(xintercept = median(output_v2$coeff_support, na.rm = TRUE),
   #          color = "red", size = 1) + 
  #theme_few() +
  #labs(x = "Generic Ballot Coefficient",
   #    y = "Count") +
  #ylim(0, 150)

#p3 <- ggplot(output_v2, aes(x = coeff_incumb)) +
#  geom_histogram(bins = 30, color = "white") 
#  geom_vline(xintercept = median(output_v2$coeff_incumb, na.rm = TRUE),
 #            color = "red", size = 1) + 
  #theme_few() +
  #labs(x = "Incumbency Coefficient",
   #    y = "Count") +
  #ylim(0, 150)

#grid.arrange(p1, p2, p3)
```

```{r, message = FALSE, echo=FALSE, warning = FALSE}
#p3
```


```{r, message = FALSE, echo=FALSE, warning = FALSE}
#summary(m)
```






