---
title: Blog Post 4
author: Kaela Ellis
date: '2022-10-03'
slug: []
categories: []
tags: []
---
```{r,  message = FALSE, echo=FALSE}
library(tidyverse)
# required packages 
require(tidyverse)
require(ggplot2)
require(sf)
library(janitor)
library(usmap)
```

Welcome to my 2022 Midterm blog series. I’m Kaela Ellis, a junior at Harvard College studying Government. This semester I am taking Gov 1347: Election Analytics taught by Professor Ryan Enos, which has led to the creation of this blog. I will use this blog series to predict the 2022 Midterm election. The series will be updated weekly with new predictive measures and analyses of past elections.

This fourth post examines the question of the effect of incumbency of elections. In particular, I look to see how adding in a factor of incumbency would affect the accuracy of my predictions. I believe that incumbency will make predictions more accurate, and should be a factor in my forecast. 

There is a widely accepted theory among forecasters called the incumbency advantage. Statistically incumbents have an advantage when it comes to reelection. I will first draw on the novel “Incumbency Advantage in US Presidential Elections: the historical record” by David Mayhew to explain the multiple reasons why elections see this incumbency advantage. Mayhew lists a few different reasons for the incumbency advantage. The first is that incumbents have capabilities while holding office that make them stronger candidates. These capabilities are things such as basic skills, having resources, ability to command attention, make pork barrel concessions, and prerogatives, such as choosing when to sign bills, make speeches, and issue executive orders at advantageous times. Another possible reason for the incumbency advantage is that incumbents are good at campaigning; they have won before and likely have a seasoned campaign manager. Another reason is that voters may be risk-averse, entering in a mentality of ‘the devil you know may be a better bet than the one you don’t know’. There are also perceived start-up costs of a new administration. There may be other factors of voter psychology, as people will revote for the incumbent to convince themselves that they did not make a mistake. Mayhew also discusses Zaller’s prizefighter theory of how incumbents are better competitors than most opponents. These are all reasons as to why during the postwar era only 4 incumbent presidents have lost reelection. 

Specific to congressional elections, incumbents who have been in office for a single term tend to lose, while incumbents who have been there for multiple terms tend to get reelected. In determining the incumbency effect on congressional elections, it may be helpful to identify the incumbents by district and determine how many sessions they have held office for. While I plan to conduct this analysis later, this week I choose to look at how expert predictions differ from actual election results. 


```{r,  message = FALSE, echo=FALSE}
expert_ratings <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 4/expert_rating.csv")
historical_results <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 4/house party vote share by district 1948-2020.csv") %>% 
  clean_names()
expert_ratings2 <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 4/2018_ratings_share (2).csv")
```

```{r,  message = FALSE, echo=FALSE}
# Selecting columns
avg_ratings <- expert_ratings %>% 
  select(year, state, district, avg_rating)  %>% filter(year == 2018)

dem_results <- historical_results %>% 
  select(race_year, state, area, dem_votes_major_percent) %>% 
  rename("year" = "race_year") %>% 
  separate(area, into = c("area", "district"), sep = " ") %>% 
  select(-area) %>% 
  mutate(district = case_when(
    district == "Large" ~ "AL",
    TRUE ~ district
  )) %>% filter(year ==2018)

```

```{r,  message = FALSE, echo=FALSE}
# Joining the data and nesting by state and district
combine_data <- avg_ratings %>% 
  # left join as there aren't ratings for every district
  left_join(dem_results, by = c("year", "state", "district"))
```

```{r,  message = FALSE, echo=FALSE}
plot_usmap(data = combine_data, regions = "states", values = "avg_rating") +
scale_fill_gradient2(high = "red", mid = "white", low = "blue",  midpoint = 4, name = "Two-party voteshare margin") +
theme_void() + labs(title = "2018 Expert Rating")

plot_usmap(data = combine_data, regions = "states", values = "dem_votes_major_percent") +
scale_fill_gradient2(high = "blue", mid = "white", low = "red",  midpoint = 50, name = "Two-party voteshare margin") +
theme_void() + labs(title = "2018 Actual Voteshare")
```
```{r, message = FALSE, echo=FALSE}
incumb <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 4/incumb_dist_1948-2022 (2).csv")

incumbwon <- incumb %>% 
select(year, winner_party, winner_candidate, winner_candidate_inc) %>%
mutate(winparty_last = lag(winner_party, order_by = year),
wincand_last = lag(winner_candidate, order_by = year),
wincand_last_inc = lag(winner_candidate_inc, order_by = year)) %>%
mutate(reelect.cand = wincand_last_inc == winner_candidate_inc) %>%
filter(year > 1948) %>%
group_by(reelect.cand) %>%
summarise(n = n()) %>%
as.data.frame() 

##how to indicate presidents party variable?
partywon <- incumb %>%
mutate(winparty_last = lag(winner_party, order_by = year),
wincand_last = lag(winner_candidate, order_by = year)) %>%
mutate(reelect.party = winparty_last) %>%
filter(year > 1948) %>%
group_by(reelect.party) %>%
summarise(n = n()) %>%
as.data.frame() 
```

