---
title: "Blog_Post_5"
author: "kaela"
date: '2022-10-11'
output: html_document
---

Welcome to my 2022 Midterm blog series. I’m Kaela Ellis, a junior at Harvard College studying Government. This semester I am taking Gov 1347: Election Analytics taught by Professor Ryan Enos, which has led to the creation of this blog. I will use this blog series to predict the 2022 Midterm election. The series will be updated weekly with new predictive measures and analyses of past elections.


Using 2018 Ads Data. Using the data from this section (and incorporating useful data
from previous weeks) fit a model and predict 2022 district-level voteshare in relevant
districts given existing data on ad spending in 2018.2. What are the limitations of your model?


```{r}
library(tidyverse)
library(ggplot2)
library(readr)
require(ggplot2)
require(sf)
library(janitor)

# load datasets
pvstate_df   <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 4/incumb_dist_1948-2022 (2).csv")
#pvstate_df   <- 'incumb_dist_1948-2020 (3)'
ad_creative  <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 5/ads_2006_2018 (1).csv")
ad_issues <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 5/ads_issues_2012-2018 (1).csv")
```



```{r}
expert_ratings <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 4/expert_rating.csv")
historical_results <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 4/house party vote share by district 1948-2020.csv") %>% 
  clean_names()
expert_ratings2 <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 4/2018_ratings_share (2).csv")
```

```{r}
# Selecting columns
avg_ratings <- expert_ratings %>% 
  select(year, state, district, avg_rating)  %>% filter(year == 2018) %>%  rename("DISTRICT" = "district") %>% rename("STATENAME" = "state") %>%  drop_na()
dem_results <- historical_results %>% 
  select(race_year, state, area, dem_votes_major_percent, district_num) %>% 
  rename("year" = "race_year") %>%  rename("DISTRICT" = "district_num") %>% rename("STATENAME" = "state") %>% filter(year ==2018)
```

```{r}
get_congress_map <- function(cong=114) {
  tmp_file <- tempfile()
  tmp_dir  <- tempdir()
  zp <- sprintf("https://cdmaps.polisci.ucla.edu/shp/districts114.zip",cong)
  download.file(zp, tmp_file)
  unzip(zipfile = tmp_file, exdir = tmp_dir)
  fpath <- paste(tmp_dir, sprintf("districtShapes/districts114.shp",cong), sep = "/")
  st_read(fpath)
}

cd114 <- get_congress_map(114)

cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
cd114 <- cd114 %>% left_join(dem_results, by=c("DISTRICT", "STATENAME"))

```

```{r}
# Selecting columns

ad_freq <- ad_creative %>%  separate(airdate, into = c("year", "month", "day"), sep = "-") %>% 
  select(year, state, district) %>% filter(year == 2018) %>%  rename("DISTRICT" = "district") %>% rename("STATENAME" = "state") 


ad_freq_pretty1 <-  ad_freq %>%  data.frame(ad_freq) %>% rename("DISTRICT" = "Var2") %>% rename("STATENAME" = "Var1") %>% rename("ad_freq" = "Freq") %>% select(STATENAME, DISTRICT, ad_freq)
  ad_freq_new <- subset(ad_freq_pretty1, ad_freq != 0) 
  ad_freq_new <- data.frame(ad_freq_new)
  
 
dem_results <- historical_results %>% 
  select(race_year, state, area, dem_votes_major_percent, district_num) %>% 
  rename("year" = "race_year") %>%  rename("DISTRICT" = "district_num") %>% rename("STATENAME" = "state") %>% filter(year ==2018)
```


```{r}

cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
ad_freq_new$DISTRICT <- as.numeric(ad_freq_new$DISTRICT)
cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
ad_freq_new$STATENAME <- as.character(ad_freq_new$STATENAME)
cd114$STATENAME <- as.character(cd114$STATENAME)

cd116 <- cd114 %>% left_join(ad_freq_new, by=c("DISTRICT", "STATENAME"))
districts_simp <- rmapshaper::ms_simplify(cd116, keep = 0.01)
ggplot() + 
  geom_sf(data=districts_simp,aes(fill=ad_freq),
          inherit.aes=FALSE,alpha=0.9) + 
  scale_fill_gradient2("Number of Ads") +
  coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 49), expand = FALSE) +  
  theme_void() + labs(title = "2018 Number of Ads by District") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```
```{r}
districts_simp <- rmapshaper::ms_simplify(cd114, keep = 0.01)
ggplot() + 
  geom_sf(data=districts_simp,aes(fill=dem_votes_major_percent),
          inherit.aes=FALSE,alpha=0.9) + 
  scale_fill_gradient2(high = "red", mid = "white", low = "blue",  midpoint = 50, name = "Two-party voteshare margin") +
  coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 49), expand = FALSE) +  
  theme_void() + labs(title = "2018 National Vote Share by Party") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
  
```

```{r}
ad_freq_hist <- ad_creative %>%  separate(airdate, into = c("year", "month", "day"), sep = "-") %>% 
  select(year, state, district, party) %>% filter(year == 2006, 2008, 2010, 2012, 2014) %>%  rename("DISTRICT" = "district") %>% rename("STATENAME" = "state") 

ad_freq_pretty2 <- data.frame(ad_freq_hist) %>% rename("DISTRICT" = "Var2") %>% rename("STATENAME" = "Var1") %>% rename("ad_freq" = "Freq") %>% select(STATENAME, DISTRICT, ad_freq, party, year)

  ad_freq_new2 <- subset(ad_freq_pretty2, ad_freq != 0) 
  ad_freq_new2 <- data.frame(ad_freq_new2)
  
 
dem_results_hist <- historical_results %>% 
  select(race_year, state, area, dem_votes_major_percent, district_num) %>% 
  rename("year" = "race_year") %>%  rename("DISTRICT" = "district_num") %>% rename("STATENAME" = "state") %>% filter(year == 2006, 2008, 2010, 2012, 2014)
```

```{r}
cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
ad_freq_new2$DISTRICT <- as.numeric(ad_freq_new2$DISTRICT)
cd114$DISTRICT <- as.numeric(cd114$DISTRICT)
ad_freq_new2$STATENAME <- as.character(ad_freq_new2$STATENAME)
cd114$STATENAME <- as.character(cd114$STATENAME)

cd117 <- cd114 %>% left_join(ad_freq_new2, by=c("DISTRICT", "STATENAME"))
districts_simp <- rmapshaper::ms_simplify(cd117, keep = 0.01)
ggplot() + 
  geom_sf(data=districts_simp,aes(fill=ad_freq),
          inherit.aes=FALSE,alpha=0.9) + 
  scale_fill_gradient2("Number of Ads") +
  coord_sf(xlim = c(-124.43, -66.57), ylim = c(23, 49), expand = FALSE) +  
  theme_void() + labs(title = "2006-2014 Number of Ads by District") +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
```

