---
title: "Blog_Post_2"
author: "kaela"
date: '2022-09-17'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Blog Post 2

Welcome to my 2022 Midterm blog series. I’m Kaela Ellis, a junior at Harvard College studying Government. This semester I am taking Gov 1347: Election Analytics taught by Professor Ryan Enos, which has led to the creation of this blog. I will use this blog series to predict the 2022 Midterm election. The series will be updated weekly with new predictive measures and analyses of past elections.

```{r,  message = FALSE, echo=FALSE}
library(tidyverse)
data <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 2/unemployment_state_monthly.csv")
# required packages 
require(tidyverse)
require(ggplot2)
require(sf)


```
Heterogenous Predictive Power of the Economy. Does the
effect of the economy vary when we consider popular vote versus
seat share as our outcome (dependent) variable? Does the predictive
power of economy change across time? If so, why?

```{r}
#use data from 7th quarter which occurs 7/1 -10/1

popvote_df <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 2/house_popvote_seats.csv") 
                                                   
# load GDP data by quarter
economy_df <- read_csv("~/Desktop/Gov1347/Gov-1347/Week 2/GDP_quarterly.csv")

#merge popvote_df and economy_df by year
dat <- left_join(economy_df, popvote_df, by = 'year')

# drop NAs (only election years)
dat <- dat %>%
  drop_na()

# new df
dat2 <- dat %>% 
    select('year', 'winner_party', 'H_incumbent_party', 'H_incumbent_party_majorvote_pct', 
           'quarter_cycle', 'GDP_growth_pct', ) %>%
    filter(quarter_cycle == 7)  

dat2 %>%
  ggplot(aes(x=GDP_growth_pct, y=H_incumbent_party_majorvote_pct,
             label=year)) + 
    geom_text() +
    geom_hline(yintercept=50, lty=2) +
    geom_vline(xintercept=0.01, lty=2) + # median
    xlab("Q7-Q8 GDP growth") +
    ylab("Incumbent party PV") +
    theme_bw() +
    theme(
      axis.text = element_text(size = 10)
    )



```
```{r}
#not include 2020
dat2.no2020 <- dat2 [-c(1, 37), ]
dat2.no2020
  dat2.no2020 %>% 
  ggplot(aes(x=GDP_growth_pct, y=H_incumbent_party_majorvote_pct,
             label=year)) + 
    geom_text() +
    geom_hline(yintercept=50, lty=2) +
    geom_vline(xintercept=0.01, lty=2) + # median
    xlab("Q7-Q8 GDP growth") +
    ylab("Incumbent party PV") +
    theme_bw() +
    theme(
      axis.text = element_text(size = 10)
    )

```
```{r}
dat2.no2020 %>%
  ggplot(aes(x=GDP_growth_pct, y=H_incumbent_party_majorvote_pct,
             label=year)) + 
    geom_text(size = 4) +
    geom_smooth(method="lm", formula = y ~ x) +
    geom_hline(yintercept=50, lty=2) +
    geom_vline(xintercept=0.01, lty=2) + # median
    xlab("Q7-Q8 GDP growth") +
    ylab("Incumbent party PV") +
    theme_bw() +
    ggtitle("Effect of Q7 GDP growth on Incumbent Party Popular Vote Share") +
  theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12),
          plot.title = element_text(size = 12))
    
```
```{r}
#We can plot the in-sample error via residuals, which capture the
#difference between each observed value (yi) and predicted value
#(ybi = Aˆ + Bx b i):
lm_econ <- lm(H_incumbent_party_majorvote_pct ~ GDP_growth_pct, 
              data = dat2.no2020)

plot(dat2.no2020$year, dat2.no2020$H_incumbent_party_majorvote_pct,
type="l",
main="True Incumbent Party PV (line), Predicted Incumbent Party PV (dot)", xlab = "Year", ylab = "Incumbent Party PV Share")
points(dat2.no2020$year, predict(lm_econ, dat2.no2020))
```

```{r}
#We can plot the in-sample error via residuals, which capture the
#difference between each observed value (yi) and predicted value
#(ybi = Aˆ + Bx b i):

hist(lm_econ$model$H_incumbent_party_majorvote_pct -
lm_econ$fitted.values,
main="In Sample Error: True PV - Predicted PV", xlab = "True Incumebent Party PV - Predicted Incumebent Party PV")
```
```{r}
#seat share as an outcome
dat2.no2020
dat2.no2020$incumbent.party.net.seats.pct <- c(-29, 22, 19, 2, 49,-21, -4, 37, -47, -5, 12, -13, 49
, 1, -15, -35, 27, -16, 5, 2, 7, -9, -54, -4, 4, -2, 8, 3, -30, 23, -63, -8, -12, -6, -42)

dat2.no2020 %>%
  ggplot(aes(x=GDP_growth_pct, y=incumbent.party.net.seats.pct,
             label=year)) + 
    geom_text(size = 4) +
    geom_smooth(method="lm", formula = y ~ x) +
    geom_hline(yintercept=0, lty=2) +
    geom_vline(xintercept=0.01, lty=2) + # median
    xlab("Q7-Q8 GDP growth (X)") +
    ylab("Incumbent party House Seat Share") +
    theme_bw() +
    ggtitle("Effect of Q7 GDP growth on Incumbent Party Seat Share") +
  theme(axis.text = element_text(size = 12),
          axis.title = element_text(size = 12),
          plot.title = element_text(size = 12))
```

```{r}
lm_econ2 <- lm(incumbent.party.net.seats.pct ~ GDP_growth_pct, 
              data = dat2.no2020)

plot(dat2.no2020$year, dat2.no2020$incumbent.party.net.seats.pct,
type="l",
main="true Y (line), predicted Y (dot) for each year")
points(dat2.no2020$year, predict(lm_econ2, dat2.no2020))
```
```{r}
hist(lm_econ2$model$incumbent.party.net.seats.pct -
lm_econ2$fitted.values,
main="In Sample Error: True Seat Share - Predicted Seat Share", xlab = "True Incumebent Party Seat Share - Predicted Incumebent Party Seat Share")
```
```{r}
outsamp_errors <- sapply(1:1000, function(i){
    years_outsamp <- sample(dat2.no2020$year, 8)
  outsamp_mod <- lm(incumbent.party.net.seats.pct ~ 
                      GDP_growth_pct,
                  dat2.no2020[!(dat2.no2020$year %in% years_outsamp),])
  outsamp_pred <- predict(outsamp_mod,
                newdata = dat2.no2020[dat2.no2020$year %in% years_outsamp,])
  outsamp_true <- dat2.no2020$incumbent.party.net.seats.pct[dat2.no2020$year 
                                                       %in% years_outsamp]
  mean(outsamp_pred - outsamp_true)
})

hist(outsamp_errors,
     xlab = "",
     main = "Seat Share Mean out-of-Sample Residual\n(1000 runs of cross-validation)")
```

